### 2020-06-19 text/article_vedrovice.Rmd
continue spatial analysis first/second-order effects

Spatial arrangement of prehistoric burial grounds is often examined to find
any internal structures, arrangements in lines, possible grouping and so forth.
We approach the task as a point pattern analysis problem given that
a burial (event) is simplified as a point location in geographic space.
The excavated polygon is used as a study area (a window).
Using a window closely approximating locations of the burials was also considered,
e. g. buffer zone or convex/concave hull but simplified excavation polygon is
deemed the most useful because it preserves the *natural* borders and empty areas
in the burial ground.

Overall structure of the burial ground is examined using both density based approach
and distance based methods [@osullivan2010, 121-154].
The density of burials is examined using quadrat test while
the distance between neighbours using the $G$, Ripley's $K$ [@ripley1976] and $L$ functions.

>>>>

Spatial arrangement of prehistoric burial grounds is often examined to find
any internal structure, arrangement or possible grouping.
We approach the task as a point pattern analysis problem given that
a burial (event) is simplified as a point location in geographic space.
A polygon of the excavation extent is used as a study area (a window).
Using a window closely approximating locations of the burials was also considered,
e. g. buffer zone or convex/concave hull but simplified excavation polygon is
deemed the most useful because it preserves the *natural* borders and empty areas
in the burial ground.

Overall structure of the burial ground is examined using both density based approach
and distance based methods [@osullivan2010, 121-154].
It is argued that density based approaches, for instance quadrat tests,
characterize the point pattern in terms of its *first-order properties* [@osullivan2010, 126].
The assumption underlying the *first-order effects* is that any point location is
independent on other point locations and different parameters influence the location
[@nakoinz2016, 130-135; @osullivan2010, 107].
On the other hand the methods based on distances between nearest neighbours
characterize the point pattern in regards to *second-order effects*.
In this case it is presumed that the point location depends on other
point locations.
Functions $G$, Ripley's $K$ [@ripley1976] and $L$ are explored as the most widely
used distance-based point pattern measures [@nakoinz2016, 135-144; @osullivan2010, 130-139]

We assume that in case of a burial ground *first-order properties* are
of little to no importance at all.
In case of spatial analysis of settlement structure, individual locations of
settlements (points) can depend on a number of parameters, for instance
elevation, soil type etc.
<!-- It is possible that the position of burials was dependent on parameters that were not
captured by the excavation, for example location of trees but that is impossible to test.-->
In case of a grave placement we presume that the choice of a location is influenced
solely by other (previous) grave locations.
The *second-order properties*, and possibly *third-order properties* [@nakoinz2016, 130, 144]
which take into account triplets of points, thus effectively the overall structure of the pattern,
are considered major factors in locating the graves.
